A prototype software framework for transparent, reusable and updatable computational health economic models

### 2023-10-22 Markdown/Manuscript/Child_RMDs/MS_Discussion.Rmd
problems with table and figure placement

However, our prototype has a number of features that future work on CHEM software frameworks may find useful to incorporate. Firstly, developing a software framework to work within an existing and widely used open source programming language such as R or Python, can keep framework scope relatively narrow. This makes it more tractable to develop, maintain and learn, and can integrate other modelling tools written in that language (e.g., the dependency libraries we list in Table \@ref(tab:cpkgs)). Secondly, implementation that combines both object-oriented and functional programming paradigms can avail of the modular and syntactical simplicity benefits of the former, while limiting needless bundling of code artefacts. Thirdly, a sensible trade-off needs to be found between transparent code implementation (which requires clear and sufficiently detailed documentation) and Agile Software Development (for which a foundational principle is prioritising the development of working code over writing documentation [@beck2001manifesto]). Our software framework makes this trade off by enforcing the use of consistent code style conventions and file organisation which in turn enables automated generation of simple documentation at every CHEM module update. All model data-structures and algorithms are therefore always documented (at least minimally, with machine authored content), meaning model developers have a requirement to write customised documentation less frequently. Many of the features of our prototype software framework have the potential to be significantly enhanced through integration with evolving code and documentation authoring capabilities of large language models.

>>>>

However, our prototype has a number of features that future work on CHEM software frameworks may find useful to incorporate. Firstly, developing a software framework to work within an existing and widely used open source programming language such as R, can keep framework scope relatively narrow. This makes it more tractable to develop, maintain and learn, and facilitates integration with other modelling tools written in that language (e.g., the dependency libraries we list in Table \@ref(tab:cpkgs)). Secondly, implementation that combines both object-oriented and functional programming paradigms can avail of the modular and syntactical simplicity benefits of the former, while limiting needless bundling of code artefacts. Thirdly, a sensible trade-off needs to be found between transparent code implementation (which requires clear and sufficiently detailed documentation) and Agile Software Development (for which a foundational principle is prioritising the development of working code over writing documentation [@beck2001manifesto]). Our software framework makes this trade off by enforcing the use of consistent code style conventions and file organisation which in turn enables automated generation of simple documentation at every CHEM module update. All model data-structures and algorithms are therefore always documented (at least minimally, with machine authored content), meaning model developers have a requirement to write customised documentation less frequently. Many of the features of our prototype software framework have the potential to be significantly enhanced through integration with evolving code and documentation authoring capabilities of large language models.

### 2023-10-22 Markdown/Manuscript/Child_RMDs/MS_Imp_Example.Rmd
problems with table and figure placement

 - programs for replicating all steps from data ingest to manuscript reporting [@hamilton_matthew_2022_6129906], applying utility mapping models to new data [@hamilton_matthew_2022_6416330] and generating a synthetic representation of the study dataset [@hamilton_matthew_p_2022_6321821]; and

>>>>

 - programs for replicating all steps from data ingest to manuscript production [@hamilton_matthew_2022_6129906], applying utility mapping models to new data [@hamilton_matthew_2022_6416330] and generating a synthetic representation of the study dataset [@hamilton_matthew_p_2022_6321821]; and

### 2023-10-22 Markdown/Manuscript/Child_RMDs/MS_Imp_Guidelines.Rmd
problems with table and figure placement

A CHEM that is **reusable** also signals ethical modelling practice. Although making a CHEM’s code, data and documentation publicly available is increasingly considered good practice, it is insufficient to enable model re-use.  Key concerns for health economists when considering model reuse are generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39]. Writing model algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task) is good scientific computing practice [@Wilson_2017] and promotes selective reuse. Computational implementations that store model code and data in distinct files and locations (as opposed to embedding data such as parameter values into source code) are easier to selectively modify. Modular implementations construct models from multiple independently reusable and replaceable sub-models (modules) [@pan2021modular]. The programming concept inheritance [@8717448], deployed in object-oriented programming approaches, can facilitate the duplication and selective modification of models. Granting permissions to others to use, test and adapt models and their components, can be facilitated by two broad categories of open source licenses. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which content can be re-used. Alternatively, copyleft licenses [@copyleft2022] can be used to require that derivative works created by content users remain open-source. 
<!-- Compared to using commercial modelling software, authoring CHEMs in an open-source language like R [@RCORE2022] makes it easier to -->
<!-- Using open-source development platforms and licenses can aid both generalizability and transferability. -->
<!-- The choices that CHEM developers make about model implementation and licensing will also shape who can use a model and for what purposes.  -->
<!-- are relatively easier to selectively modify. -->

@@

##	Assessment criteria
For each CHEM attribute that suggests ethical CHEM modelling practice, we identified assessment criteria that have the potential to be met with the assistance of software frameworks.

@@

-	**U2**: Each new release of a model is systematically retested, with changes implemented to minimise disruptions for existing model users.

>>>>

A CHEM that is **reusable** also signals ethical modelling practice. Although making a CHEM’s code, data and documentation publicly available is increasingly considered good practice, it is insufficient to enable model re-use.  Key concerns for health economists when considering whether to reuse a model are generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39]. Writing model algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task) is good scientific computing practice [@Wilson_2017] and promotes selective reuse. Computational implementations that store model code and data in distinct files and locations (as opposed to embedding data such as parameter values into source code) are easier to selectively modify. Modular implementations enable models to be constructed from multiple independently reusable and replaceable sub-models (modules) [@pan2021modular]. The programming concept inheritance [@8717448], deployed in object-oriented programming approaches, can facilitate the duplication and selective modification of models. Granting permissions to others to use, test and adapt models and their components, can be facilitated by two broad categories of open source licenses. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which content can be re-used. Alternatively, copyleft licenses [@copyleft2022] can be used to require that derivative works created by content users remain open-source. 
<!-- Compared to using commercial modelling software, authoring CHEMs in an open-source language like R [@RCORE2022] makes it easier to -->
<!-- Using open-source development platforms and licenses can aid both generalizability and transferability. -->
<!-- The choices that CHEM developers make about model implementation and licensing will also shape who can use a model and for what purposes.  -->
<!-- are relatively easier to selectively modify. -->

@@

##	Assessment criteria
For each CHEM attribute that suggests ethical CHEM modelling practice, we identified assessment criteria. <!-- that have the potential to be met with the assistance of software frameworks -->

@@

-	**U2**: Each new release of a CHEM is systematically retested, with changes implemented to minimise disruptions for existing model users.

### 2023-10-22 Markdown/Manuscript/Child_RMDs/MS_Tables_2.Rmd
problems with table and figure placement

df <- data.frame(`CHEM attribute` = c("Transparent", "Reusable", "Updatable"),

@@

                 `Beneficial use` = c("Decision makers more likely to understand strengths and weakness of model before applying it.","Can be used to inform more decision problems, in more contexts and can help reduce duplicative mode development effort.","Models that are maintained can be validly used for longer."),

@@

    flextable::flextable(df) %>%
      flextable::add_header_row(values = c(" ","Project team responsibility"),colwidths = c(1,3)) %>%
      flextable::border_inner(border = officer::fp_border(style = "none")) %>%
      flextable::border_outer() %>%
      flextable::padding(padding.top = 5, part = "all") %>%
      flextable::padding(padding.bottom = 5, part = "all") %>%
            flextable::padding(padding.left = 5, part = "all") %>%
      flextable::padding(padding.right = 5, part = "all") %>%
      flextable::align(align = "center", part = "header") %>%
      flextable::merge_h() %>%
      flextable::merge_v() %>%
      flextable::bold(j=1, part = "header") %>%
      flextable::width(width = 1.5)

@@

    # df %>%
    #   kableExtra::kbl(booktabs = T, 
    #                   escape = F,
    #                   caption = knitr::opts_current$get("tab.cap")) %>%
    #   kableExtra::kable_paper(full_width = F) %>%
    #   kableExtra::column_spec(1, bold=T) %>%
    #   kableExtra::add_header_above(c(" ", "Project team responsibility" = 3), bold = T, italic = T) %>%
    #   kableExtra::column_spec(2, width = "13em") %>%
    #   kableExtra::column_spec(3, width = "13em") %>%
    #   kableExtra::column_spec(4, width = "13em") %>%
    #   kableExtra::collapse_rows(columns = 2, latex_hline = "none")

>>>>

df <- data.frame(` ` = c("Transparent", "Reusable", "Updatable"),

@@

                 `Beneficial use` = c("Decision-makers can understand the strengths and weakness of models before applying them.","Can inform more decision problems, in more contexts with less duplicative effort.","Models can be validly used for longer."),

@@

    # flextable::flextable(df) %>%
    #   flextable::border_inner(border = officer::fp_border(style = "none")) %>%
    #   flextable::border_outer() %>%
    #   flextable::align(align = "center", part = "header") %>%
    #   flextable::merge_h() %>%
    #   flextable::merge_v() %>%
    #   flextable::valign(j=2, valign = "top") %>%
    #   flextable::align(j=2:4, align = "center") %>%
    #   flextable::bold(j=1, part = "all") %>%
    #   flextable::bold(part = "header") %>%
    #   flextable::width(j = 1, width = 1.2) %>%
    #   flextable::width(j = 2:4, width = 1.6)

@@

      kableExtra::kbl(booktabs = T,
                      escape = F,
                      caption = knitr::opts_current$get("tab.cap")) %>%
      kableExtra::kable_paper(full_width = F) %>%
      kableExtra::column_spec(1, bold=T) %>%
      kableExtra::add_header_above(c(" ", "Project team responsibility" = 3), bold = T, italic = T) %>%
      kableExtra::column_spec(2, width = "13em") %>%
      kableExtra::column_spec(3, width = "13em") %>%
      kableExtra::column_spec(4, width = "13em") %>%
      kableExtra::collapse_rows(columns = 2, latex_hline = "none")

### 2023-10-22 Markdown/Manuscript/Child_RMDs/MS_Tables_3.Rmd
problems with table and figure placement

names_chr <- names(df)
# df <- data.frame(df[,1],
#                  rep(ifelse(params$X@outp_formats_chr[1] == "Word"," ","&nbsp; &nbsp; &nbsp;"),
#                      nrow(df)),
#                  df[,2#:3
#                     ],
#                  rep(ifelse(params$X@outp_formats_chr[1] == "Word"," ","&nbsp; &nbsp; &nbsp;")
#                      ,nrow(df)),
#                  #df[,4],
#                  #rep("&nbsp; &nbsp; &nbsp;",nrow(df)),
#                  df[,3#5
#                     ])
# colnames(df) <- c("Package",
#                   ifelse(params$X@outp_formats_chr[1] == "Word","XXX"," "),
#                   #"Ref", 
#                   "Purpose", 
#                   ifelse(params$X@outp_formats_chr[1] == "Word","YYY"," "),
#                   #"Standard     ",# " ",
#                   "R library dependencies")
#df$Package <- "![](../Data/images/ready4.png){ width=10% }"#"\\includegraphics[valign=T,scale=2.5,raise=2mm]{../Data/images/ready4.png}"
#df$`Standard     `  <- df$`Standard     ` %>% stringr::str_replace_all("&nbsp;"," ")
if(params$X@outp_formats_chr[1] == "Word") {
  library("flextable")
  library("ftExtra")
  df[,c(1,3:4,6:7)] %>%
    ready4show::print_table(output_type_1L_chr = params$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL) %>%
    ftExtra::colformat_md()
  }else{
    #df$Ref<-NULL
    #df$Package <- ""
    df$Package  <- df$Package %>% 
                      purrr::map_chr(~paste0("\\includegraphics[valign=M,width=1.9cm,raise=4mm]{../Data/images/",.x,"_logo.png}"))#scale=0.4,raise=3mm]
    # df$Package <- "\\includegraphics[valign=M,scale=0.08,raise=3mm]{../Data/images/ready4.png}"
    #df[,c(1,3,5:6)]   %>%
    df %>%
    # kableExtra::kbl(format = "markdown",
    #                   escape = F,
    #                 caption = knitr::opts_current$get("tab.cap"))
      kableExtra::kbl(booktabs = T, 
                      escape = F,
                      caption = knitr::opts_current$get("tab.cap")) %>%
      kableExtra::kable_paper(full_width = F) %>%
      kableExtra::column_spec(2, width = "20em", latex_valign = "m") %>%
      # kableExtra::column_spec(1, 
      #                         image = kableExtra::spec_image(rep("../Data/images/ready4.png",nrow(df)), 
      #                                                           150, 150)) %>%
      kableExtra::column_spec(3, width = "20em",
                              latex_valign = "m")
    }
```

>>>>

names_chr <- names(df)
# df <- data.frame(df[,1],
#                  rep(ifelse(params$X@outp_formats_chr[1] == "Word"," ","&nbsp; &nbsp; &nbsp;"),
#                      nrow(df)),
#                  df[,2#:3
#                     ],
#                  rep(ifelse(params$X@outp_formats_chr[1] == "Word"," ","&nbsp; &nbsp; &nbsp;")
#                      ,nrow(df)),
#                  #df[,4],
#                  #rep("&nbsp; &nbsp; &nbsp;",nrow(df)),
#                  df[,3#5
#                     ])
# colnames(df) <- c("Package",
#                   ifelse(params$X@outp_formats_chr[1] == "Word","XXX"," "),
#                   #"Ref", 
#                   "Purpose", 
#                   ifelse(params$X@outp_formats_chr[1] == "Word","YYY"," "),
#                   #"Standard     ",# " ",
#                   "R library dependencies")
#df$Package <- "![](../Data/images/ready4.png){ width=10% }"#"\\includegraphics[valign=T,scale=2.5,raise=2mm]{../Data/images/ready4.png}"
#df$`Standard     `  <- df$`Standard     ` %>% stringr::str_replace_all("&nbsp;"," ")
if(params$X@outp_formats_chr[1] == "Word") {
  library("flextable")
  library("ftExtra")
  df %>% #[,c(1,3:4,6:7)] %>%
    ready4show::print_table(output_type_1L_chr = params$X@outp_formats_chr[1],
                            caption_1L_chr = knitr::opts_current$get("tab.cap"),
                            mkdn_tbl_ref_1L_chr = paste0("tab:",knitr::opts_current$get("tab.id")),
                            use_rdocx_1L_lgl = ifelse(params$X@outp_formats_chr[1]=="Word",T,F),
                            add_to_row_ls = NULL,
                            sanitize_fn = NULL) %>%
    ftExtra::colformat_md()
  }else{
    #df$Ref<-NULL
    #df$Package <- ""
    # df$Package  <- df$Package %>% 
    #                   purrr::map_chr(~paste0("\\includegraphics[valign=M,width=1.9cm,raise=4mm]{../Data/images/",.x,"_logo.png}"))#scale=0.4,raise=3mm]
    # df$Package <- "\\includegraphics[valign=M,scale=0.08,raise=3mm]{../Data/images/ready4.png}"
    #df[,c(1,3,5:6)]   %>%
    df %>%
    # kableExtra::kbl(format = "markdown",
    #                   escape = F,
    #                 caption = knitr::opts_current$get("tab.cap"))
      kableExtra::kbl(booktabs = T, 
                      escape = F,
                      caption = knitr::opts_current$get("tab.cap")) %>%
      kableExtra::kable_paper(full_width = F) %>%
      kableExtra::column_spec(2, width = "20em", latex_valign = "m") %>%
      # kableExtra::column_spec(1, 
      #                         image = kableExtra::spec_image(rep("../Data/images/ready4.png",nrow(df)), 
      #                                                           150, 150)) %>%
      kableExtra::column_spec(3, width = "20em",
                              latex_valign = "m")
    }
```

### 2023-10-22 Markdown/Manuscript/Child_RMDs/MS_Tables_4.Rmd
problems with table and figure placement

                               "T2 Judgments and contributions are easily identified", 

@@

                 Detail = c("All source code and testing procedures are available in public GitHub repositories, with each code release persistently available in a Zenodo repository. Study dataset contains confidential patient records and was not published. Instead, a synthetic representation of the study dataset is persistently available in a repository in the Harvard Dataverse. Data files to support out of sample application of models are published at the same location.",
                            "All module libraries, programs, and datasets are distributed with citation information. GitHub repositories detail author code contributions over project development history are publicly visible. Model catalogues persistently available on the Harvard Dataverse describe the predictive performance of models under multiple usage regimes. Each code library is documented with worked examples of how to apply modules.  Analysis and reporting programs are self-documenting.  Sub-routines for generating reports are documented with README files.",
                            "Model code is written using both functional and object-oriented paradigms. Code library websites include hypothetical examples of generalisability (applying study algorithm to estimate mapping models from new data with the same predictor and outcome variables) and transferability (adapting study algorithm to develop mapping models from datasets with different predictor variables and outcomes measured with a different utility instrument)." ,
                            "All code is distributed using GPL-3 licenses. Datasets use amended version of template provided by Harvard Dataverse, allowing reuse of data subject to some ethical restrictions (e.g., use in efforts to reidentify study participants is prohibited).",

@@

                            "Continuous integration used for all code libraries, primarily for acceptance testing. Only limited use is made of unit testing. Retired library code is deprecated using tools from the lifecycle R library. Library documentation articles and datasets are also deprecated."),

>>>>

                               "T2 Project team contributions and judgments are easily identified", 

@@

                 Detail = c("All source code and testing procedures are available in public GitHub repositories, with each code release persistently available in a Zenodo repository. As the study dataset contains confidential patient records, it was not published. Instead, a synthetic representation of the study dataset is persistently available in a repository in the Harvard Dataverse. Data files to support out of sample application of models are published at the same location.",
                            "All module libraries, programs, and datasets are distributed with citation information. Public GitHub repositories detail author code contributions over project development history. Model catalogues persistently available on the Harvard Dataverse describe the predictive performance of models under multiple usage regimes. Each code library is documented with worked examples of how to apply CHEM modules.  Analysis and reporting programs are self-documenting.  Sub-routines for generating reports are documented with README files.",
                            "CHEMs are written using both functional and object-oriented paradigms. Code library documentation websites include hypothetical examples of generalisability (applying study algorithm to estimate mapping models from new data with the same predictor and outcome variables) and transferability (adapting study algorithm to develop mapping models from datasets with different predictor variables and outcomes measured with a different utility instrument)." ,
                            "All code is distributed using GPL-3 licenses. Datasets use an amended version of template terms provided by Harvard Dataverse, allowing reuse of data subject to some ethical restrictions (e.g., use in efforts to reidentify study participants is prohibited).",

@@

                            "Continuous integration is used for all code libraries, primarily for acceptance testing. Only limited use is made of unit testing. Retired library code is deprecated using tools from the lifecycle R library. Library documentation articles and datasets are also deprecated."),

### 2023-10-22 Markdown/Manuscript/Parent_PDF/Main_Bookdown_PDF.Rmd
problems with table and figure placement

---
bibliography: ../Data/references.bib
output: 
  bookdown::pdf_document2: 
    includes:
      in_header: preamble.tex
    pandoc_args: 
      - "--metadata-file=../Header/header_common.yaml"
      - '--lua-filter=../Filters/scholarly-metadata.lua'
      - '--lua-filter=../Filters/author-info-blocks.lua'
    toc: no
params:
  X: NULL
abstract: "`r paste0(readLines('../Header/abstract.txt'),collapse=' ')`"
graphics: yes
csl: plos.csl
---

>>>>

---
bibliography: ../Data/references.bib
output: 
  bookdown::pdf_document2: 
    toc: no
    keep_tex: true
    includes:
      in_header: preamble.tex
    pandoc_args: 
      - "--metadata-file=../Header/header_common.yaml"
      - '--lua-filter=../Filters/scholarly-metadata.lua'
      - '--lua-filter=../Filters/author-info-blocks.lua'
params:
  X: NULL
abstract: "`r paste0(readLines('../Header/abstract.txt'),collapse=' ')`"
graphics: yes
csl: plos.csl
---
